---
title: "Homework 8"
output: html_notebook
---

*Question 1*
Sum the total population (TotPop) for each Congressional District. The District numbers for the 113th Congress are identified by the variable cd113. Use the functions mean() and sd() to calculate the mean and standard deviation of the District populations. Is the standard deviation small or large relative to the mean?

```{r}
library(tidyverse)
q1 <- fl %>%
  group_by(cd113) %>%
  summarize(DistrictPop = sum(TotPop))

mean(q1$DistrictPop)
sd(q1$DistrictPop)

#mean:696344.8
#sd:178.3987

#The standard deviation is relatively small compared to the mean.

```

*Question 2*
Now sum the voting-age population (VAP) for each Congressional District. Calculate the mean and standard deviation of the District voting-age populations. What do you notice about the difference between this standard deviation and the standard deviation you calculated above? What does it say about “one person one vote?”

```{r}
q2 <- fl %>%
  group_by(cd113) %>%
  summarize(VAPpop = sum(VAP))

mean(q2$VAPpop)
sd(q2$VAPpop)

#mean: The standard deviation is relatively small compared to the mean.
#sd:17382.9

#The standard deviation is relatively large compared to the mean. Some congressional districts have a greater or lesser weight than others when it comes to how much power is in one vote.
```

*Question 3*
What percent of Florida’s total population is Black? What percent of Florida’s 27 Congressional Districts have a constituency that is a majority Black? What do you make of the difference? (Include all ages.)

```{r}
q3 <- fl %>%
  summarize(totalpop = sum(TotPop),
            totalblack = sum(BlackPop)) %>%
  mutate(percent = totalblack/totalpop *100)

q3$percent

#

q3 <- fl %>%
  group_by(cd113) %>%
  summarize(totalpop = sum(TotPop),
            totalblack = sum(BlackPop)) %>%
  mutate(percent = totalblack/totalpop *100)

length(q3$percent[q3$percent > 50])/27 *100

#11.11111

# Florida is 16 percent black but only 11.1 percent of congressional districts are majority black. 

```


*Question 4*
What percent of Florida’s total population is Hispanic? What percent of Florida’s 27 Congressional Districts have a constituency that is a majority Hispanic? What do you make of the difference? (Include all ages.)

```{r}

q4 <- fl %>%
  summarize(totalpop = sum(TotPop),
            totalhisp = sum(HispPop)) %>%
  mutate(percent = totalhisp/totalpop *100)

q4$percent

q4 <- fl %>%
  group_by(cd113) %>%
  summarize(totalpop = sum(TotPop),
            totalhisp = sum(HispPop)) %>%
  mutate(percent = totalhisp/totalpop *100)

length(q4$percent[q4$percent > 50])/27 *100

# Florida is 22.5 percent Hispanic but only 11.1 percent of congressional districts are majority black. Hispanics in Florida are have the same congressional representation as blacks, despite making up a larger percentage of the population.  


```


*Question 5*
What percent of the two-party vote did McCain receive in Florida? In what percent of Congressional Districts did McCain win a majority of the two-party vote? What do you make of the difference?

```{r}

q5<-fl%>%
  summarise(mccainsum = sum(mccain),
            obamasum = sum(obama)) %>%
  mutate(mccainvote = mccainsum/(obamasum+mccainsum))

q5$mccainvote *100

q5<-fl%>%
  group_by(cd113) %>%
  summarise(mccainsum = sum(mccain),
            obamasum = sum(obama)) %>%
  mutate(mccainvote = mccainsum/(obamasum+mccainsum) *100)

length(q5$mccainvote[q5$mccainvote > 50])/27 *100

#McCain received 48.6 percent of the two-party vote in FL, but won 63 percent of districts. Summed over all residents (in a direct democracy), McCain would have missed a majority to win the state of Florida but with the congressional district system, he was able to win in enough of them to tip the State towards him. 


```


*Question 6*
For all units with a voting-age population greater than 200, make a scatter plot that displays the relationship between the black share of the voting age population (on the x-axis) and the McCain share of the two-party vote (on the y-axis). Do the same for Hispanics. Use geom_point() to create the points of the scatter plot. Add geom_smooth() to plot the trend (think of this as a rolling average). What do you make of the relationship?


```{r}
q6 <- fl %>%
  filter(VAP > 200) %>%
  mutate(blackshare = BlackVAP/(VAP), 
         McCainShare = mccain/(mccain+obama))

ggplot(data = q6, aes(x=blackshare,y=McCainShare)) + geom_point() + geom_smooth()+theme_bw() + xlab("Black Share of the Voting Age Population") + ylab("McCain Share of the Vote")


#For blacks, as the percentage of blacks goes up in a unit, the unit is less likely to support McCain.

```

```{r}
q6 <- fl %>%
  filter(VAP > 200) %>%
  mutate(hispshare = HispVAP/(VAP), 
         McCainShare = mccain/(mccain+obama))

ggplot(data = q6, aes(x=hispshare,y=McCainShare)) + geom_point() + geom_smooth() +theme_bw()+ xlab("Hispanic Share of the Voting Age Population") + ylab("McCain Share of the Vote")

#For hispanics, we see a similar trend except past 50 percent they tend to vote more for McCain. Likely republican cubans.
```


*Question 7*
Do the same as you did for 6, except plot voter turnout on the y-axis instead. To do this, use (mccain + obama)/VAP as a measure for turnout. (Note: There are some units where the presidential vote-share is greater than the voting-age population. This is due to estimation errors in the data. Ignore these cases by constraining the y-axis to only contain values between 0 and 1.) What do you make of the relationship?

```{r}
q7 <- fl %>%
  filter(VAP > 200) %>%
  mutate(blackshare = BlackVAP/(VAP), 
         turnout = (mccain + obama)/VAP)

ggplot(data = q7, aes(x=blackshare,y=turnout)) + geom_point() + geom_smooth() + scale_y_continuous(breaks= seq(0,1,.1), limits = c(0,1) ) +theme_bw() + ylab("Turnout") + xlab("Black Share of the Vote")
```

```{r}

q7 <- fl %>%
  filter(VAP > 200) %>%
  mutate(hispshare = HispVAP/(VAP), 
          turnout = (mccain + obama)/VAP)

ggplot(data = q7, aes(x=hispshare,y=turnout)) + geom_point() + geom_smooth() +theme_bw() + scale_y_continuous(breaks= seq(0,1,.1), limits = c(0,1) ) + ylab("Turnout") + xlab("Hispanic Share of the Vote")

#As the black proportion of blocks increases, the turnout increases. The trend is opposite for hispanics. 
```


*Question 8*

Calculate McCain’s share of the two-party vote for each of Florida’s 27 Congressional District and plot the results. Arrange the Districts along the x-axis by their degree of support for McCain (from least to most). Then plot McCain’s vote share along the y-axis. Your plot should look like the one below. What do you notice about the distribution of the partisan vote?

```{r}

q8 <- fl %>%
  group_by(cd113) %>%
  summarize(McCainShare =  as.numeric(sum(mccain)/(sum(mccain) + sum(obama))))%>%
  arrange(McCainShare)

ggplot(data = q8, aes(x=reorder(cd113, q8$McCainShare), y = McCainShare)) + geom_point() + theme_bw()+ scale_y_continuous(breaks= seq(0,1,.05), limits = c(0,1)) + geom_hline(yintercept = .50) + ylab("John McCain's Share of the 2−Party Vote") + xlab("Congressional Districts Arranged by Support for McCain")

#There are more districts that swing republican despite the fact that overall in the state, there are more democrats (looks like packing!).
```



*Question 9*
Take a look at the dataframe sims. Each column in the dataframe is the result of a computer simulated redistricting procedure that assigns each observation in the fl dataset to one of the 27 Congressional Districts. Like Florida’s actual Congressional Districts, these districts are drawn to be equally apportioned, compact, and contiguous. However, unlike Florida’s Congressional Districts, these Districts are NOT drawn by politicians who have the incentive to gerrymander. Instead, they are drawn by a computer, without partisan intent. 

Your task is to calculate the McCain vote-share for every district in each of the 100 simulations. Then, for each simulation, count the number of districts where McCain has a majority of the two-party vote. Report the mean and standard deviation. Does the actual number of McCain-Majority districts fall within 2 standard deviations of the simulated mean?

(NOTE: sims has 100 columns corresponding to the results of 100 simulations. In each column there is a set of district numbers that are assigned to the 15,640 units in the dataframe fl. The rows of sims directly map to the rows of fl such that the first row of sims corresponds to the first row of fl, the second row of sims corresponds to the second row of fl, and so on.)


```{r}

test <- cbind(last, sims)

test <- test %>%
  gather("simulation","district",3:102)%>%
  group_by(simulation, district) %>%
  summarize(mccainshare = sum(mccain)/(sum(mccain) + sum(obama)) * 100)%>%
  mutate(districtwon = ifelse(mccainshare > 50, 1, 0)) %>%
  group_by(simulation) %>%
  summarise(districtswon = sum(districtwon))

mean(test$districtswon) 
#13.7
sd(test$districtswon)
#1.226805

mean(test$districtswon)/27
#0.5074074
sd(test$districtswon)/27
#0.04543722

#In real life (or what we found in question 5), McCain won 63 percent of all the districts. Using 100 simulation, we find that that result (63 percent) does not fall eithin 2 st.dvs of the simulation mean, implying there was district drawing based on partisan motivations.

#The same can be argued by looking at the district mean. Within the simulations, we found he would win 13.7 districts with a 1.2 standard deviation. In real life he won 17. 13.7 + 2(1.2) < 17

```

